Nonparametric Goodness-of-Fit Tests \\ for Discrete Null Distributions

### 2010-10-12 JSS/article.Rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@16 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

%% an abstract and keywords
\Abstract{
   A general theory for extending nonparametric goodness-of-fit tests to discrete null distributions has existed
   for several decades. Despite this, modern statistical software have generally failed to provide this methodology
   to users. The packages \pkg{ks.test} and \pkg{cvm.test} serve to fill this need in the \proglang{R} language for
   the two most popular nonparametric tests. The structure of these two packages are explained and examples
   to their specific usage are presented. Particular attention is given to the various numerical issues that arise
   in their implementation.
}
\Keywords{Cramer von Mises, goodness-of-fit tests, Kolmogorov-Smirnov, \proglang{R}}
\Plainkeywords{Cramer von Mises, goodness-of-fit tests, Kolmogorov-Smirnov, R} %% without formatting
%% at least one keyword must be supplied

@@

<<setup,include=FALSE,echo=FALSE>>=
options(keep.source = TRUE, width = 60)
@

@@

@@

@@

@@

%Discuss the discrete version. How to calculate W2, U2, and A2. 
Generalizations of the Cramer-von Mises tests were developed in \cite{choulakian1994}. Much like for the Kolmogorov-Smirnov test,
the theoretical form of the test statistics are unchanged; although the discreteness allows for a slightly simpler representation.
The null distribution of the test statistics are again distribution dependent, unlike the continuous version. The methods do not
suggest finite sample results, but rather show that the asymptotic null distribution is equal to a weighted sum of independent
chi-squared variables (the weights depending on the particular distribution). This asymptotic distribution is what we implement here;
the original papers shows that this approximation

@@

\section{Kolmogorov-Smirnov and Cramer von Mises in R}

For the user, the function \pkg{cvm.test} was designed to work in exactly the same manner as \pkg{ks.test} in \proglang{R}. This
alleviates the need to learn a new set of input and output instructions and hopefully will increase the usage of these tests. 
As there is no default functionality in \proglang{R} for continuous Cramer-von Mises, subsequent versions of the package
should also originally implement the continuous case even though this is not what we concentrate on here.

\section{Examples}

@@

Issues revolving around incorrect usage of default ks.test() in R because of continuity error (i.e. ks.test(1:2, ecdf(1:2))

Further discuss numerical issues; ks -> n over 30 has problems, CvM -> integrate function can sometimes refuse to run because of failure to calculate integral.

@@

Possibly go into a brief discussion of classes and class inheritance in R. The use of `htest' in our functions; benefits and issues with using it. Particularly troublesome in ks.test where we have a range of p-values.

@@

\section{Examples}

>>>>

%% an abstract and keywords
\Abstract{
   A general theory for extending nonparametric goodness-of-fit tests to discrete null distributions has existed
   for several decades. Despite this, modern statistical software have generally failed to provide this methodology
   to users. The packages \pkg{ks.test} and \pkg{cvm.test} serve to fill this need in the \proglang{R} language for
   the two most popular nonparametric tests. The structure of these two packages are explained and examples
   to their specific usage are presented. Particular attention is given to the various numerical issues that arise
   in their implementation.

*** try to clean up, avoid package/function issue, make it a general
contribution if possible.
}
\Keywords{Cramer von Mises, goodness-of-fit tests, Kolmogorov-Smirnov, \proglang{R}}
\Plainkeywords{Cramer von Mises, goodness-of-fit tests, Kolmogorov-Smirnov, R} %% without formatting
%% at least one keyword must be supplied

@@

<<setup,include=FALSE,echo=FALSE>>=
options(keep.source = TRUE, width = 60)
set.seed(1)
@

@@

** Ditto on wordsmithing the contribution (package/function, etc...).

@@

*** Check Cramer-von Mises for hyphen, and let's get the accent done properly.

@@

*** New paragraph, perhaps here, on the differences between these three tests.

@@

%Discuss the discrete version. How to calculate W2, U2, and A2. 
Generalizations of the Cramer-von Mises tests were developed in \cite{choulakian1994}. Much like for the Kolmogorov-Smirnov test,
the theoretical form of the test statistics are unchanged; although the discreteness allows for a slightly simpler representation.
The null distribution of the test statistics are again distribution dependent, unlike the continuous version. The methods do not
suggest finite sample results, but rather show that the asymptotic null distribution is equal to a weighted sum of independent
chi-squared variables (the weights depending on the particular distribution). This asymptotic distribution is what we implement here;
the original papers shows that this approximation (*** NOT FINISHED?)

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Kolmogorov-Smirnov and Cramer von Mises Tests in R}

Package \pkg{ks.test} contains function \code{ks.test()}, a revision of
\proglang{R}'s Kolmogorov-Smirnov function \code{ks.test()} from base
package \pkg{stats}.
Package \pkg{cvm.test} contains function \code{cvm.test()}, a new
function for Cramer von Mises tests.  Both are available from the authors
as of October, 2010,
or from R-Forge (\url{https://r-forge.r-project.org/R/?group_id=802});
they will be proposed for inclusion in \pkg{stats} in late 2010.

The revised \pkg{ks.test()} function supports one-sample tests for discrete
null distributions by allowing the second argument, \code{y}, to be
an empirical cumulative distribution function (an \proglang{R} function
with class \code{ecdf}) or an object of class \code{stepfun} specifying
a discrete distribution.  As in the original version of \code{ks.test()},
the presence of ties in the data (the first argument, \code{x}) generates a
warning unless \code{y} describes a discrete distribution.  When a discrete
distribution is specified, exact p-values are not available for 
two-sided alternative hypotheses, but the reported p-values will be
conservative.  For one-sided tests,
exact p-values are calculated using Conover's
method (when \code{exact = NULL} or \code{exact = TRUE})
if the sample size is less than or equal to 30; otherwise, asymptotic distributions
are used which are reliable in such cases (CITATION? Is this correct?).
When \code{exact = FALSE} the asymptotic distribution is used which is known
to be imprecise but conservative, even for small sample (CITATION).

*** Discussion of what we might have broken: cases where the user
provided a discrete distribution to the original ks.test() even though
it wasn't intended to be supported. ***

The function \pkg{cvm.test()} is similar to \pkg{ks.test()}.  Its first two
arguments specify the data and null distribution; the only extra option,
\code{type}, specifies the variant of the Cramer-von Mises test:
\begin{itemize}
\item \code{x}: a numerical vector of data values.
\item \code{y}: an \code{ecdf} or step-function (\code{stepfun}) for specifying
the null model
\item \code{type}: the variant of the Cramer-von Mises test; \code{W2}
is the default and most common method, \code{U2} is for cyclical data,
and \code{A2} is the Anderson-Darling alternative.
\end{itemize}
As with \code{ks.test()}, \code{cvm.test()} returns an object of class 
\code{htest}.

%%%%%%%%%%%%%%%%%%
\section{Examples}

Hmm... I thought the package had a trivial example where the standard
\code{ks.test()} was a problem?
<<ks.test>>=
library(ks.test)
x <- sample(1:10, 25, replace=TRUE)
x
ks.test(x, ecdf(1:10))
ks.test(x, ecdf(1:10), alternative="g")
ks.test(x, ecdf(1:10), alternative="g", exact=FALSE)
@

Discuss this, nice example.  Relate to new material that is about to go into
the earlier section.
<<>>=
stats::ks.test(c(0,1), ecdf(c(0,1)))
ks.test(c(0,1), ecdf(c(0,1)))
@

JAY: simulation showing original ks.test() was conservative, Conover exact.

<<cvm.test>>=
library(cvm.test)
cvm.test(x, ecdf(1:10))
cvm.test(x, ecdf(1:10), type='A2')
@

TAYLOR: good cyclical example, relating to new material about to go into
earlier section.

@@

Issues revolving around incorrect usage of default ks.test() in R because of continuity error (i.e. ks.test(1:2, ecdf(1:2)).  Will be earlier, needed here?

Further discuss numerical issues; ks -> n over 30 has problems, CvM -> integrate function can sometimes refuse to run because of failure to calculate integral.
Earlier, needed here?

@@

Possibly go into a brief discussion of classes and class inheritance in R. The use of `htest' in our functions; benefits and issues with using it. Particularly troublesome in ks.test where we have a range of p-values.  More precisely, explain that we
could provide upper and lower bounds for the p-value but are limited by class
\code{htest}.

@@

Future work: continuous CVM.

*** Bibliography needs cleaning before we finish up, just don't forget.

### 2010-10-12 JSS/article.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@16 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\subsection{Algorithm}

@@

For the user, the function \pkg{cvm.test} was designed to work in exactly the same manner as \pkg{ks.test} in \proglang{R}. This
alleviates the need to learn a new set of input and output instructions and hopefully will increase the usage of these tests. 
As there is no default functionality in \proglang{R} for continuous Cramer-von Mises, subsequent versions of the package
should also originally implement the continuous case even though this is not what we concentrate on here.

@@

@@

Go into more detail about reasons for using these versus chi-squared type tests.

Issues revolving around incorrect usage of default ks.test() in R because of continuity error (i.e. ks.test(1:2, ecdf(1:2))

Further discuss numerical issues; ks -> n over 30 has problems, CvM -> integrate function can sometimes refuse to run because of failure to calculate integral.

Benefits and drawbacks of using simulation to gather p-values.

Possibly go into a brief discussion of classes and class inheritance in R. The use of `htest' in our functions; benefits and issues with using it. Particularly troublesome in ks.test where we have a range of p-values.

Two-sample cases; no theory for discrete distributions. Why? Distribution dependence messes things up. Discuss this for sure.

>>>>

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Kolmogorov-Smirnov and Cramer von Mises Tests in R}

Package \pkg{ks.test} contains function \code{ks.test()}, a revision of
\proglang{R}'s Kolmogorov-Smirnov function \code{ks.test()} from base
package \pkg{stats}.
Package \pkg{cvm.test} contains function \code{cvm.test()}, a new
function for Cramer von Mises tests.  Both are available from the authors
as of October, 2010,
or from R-Forge (\url{https://r-forge.r-project.org/R/?group_id=802});
they will be proposed for inclusion in \pkg{stats} in late 2010.

The revised \pkg{ks.test()} function supports one-sample tests for discrete
null distributions by allowing the second argument, \code{y}, to be
an empirical cumulative distribution function (an \proglang{R} function
with class \code{ecdf}) or an object of class \code{stepfun} specifying
a discrete distribution.  As in the original version of \code{ks.test()},
the presence of ties in the data (the first argument, \code{x}) generates a
warning unless \code{y} describes a discrete distribution.  When a discrete
distribution is specified, exact p-values are not available for 
two-sided alternative hypotheses, but the reported p-values will be
conservative.  For one-sided tests,
exact p-values are calculated using Conover's
method (when \code{exact = NULL} or \code{exact = TRUE})
if the sample size is less than or equal to 30; otherwise, asymptotic distributions
are used which are reliable in such cases (CITATION? Is this correct?).
When \code{exact = FALSE} the asymptotic distribution is used which is known
to be imprecise but conservative, even for small sample (CITATION).

*** Discussion of what we might have broken: cases where the user
provided a discrete distribution to the original ks.test() even though
it wasn't intended to be supported. ***

The function \pkg{cvm.test()} is similar to \pkg{ks.test()}.  Its first two
arguments specify the data and null distribution; the only extra option,
\code{type}, specifies the variant of the Cramer-von Mises test:
\begin{itemize}
\item \code{x}: a numerical vector of data values.
\item \code{y}: an \code{ecdf} or step-function (\code{stepfun}) for specifying
the null model
\item \code{type}: the variant of the Cramer-von Mises test; \code{W2}
is the default and most common method, \code{U2} is for cyclical data,
and \code{A2} is the Anderson-Darling alternative.
\end{itemize}
As with \code{ks.test()}, \code{cvm.test()} returns an object of class 
\code{htest}.

%%%%%%%%%%%%%%%%%%
\section{Examples}

Hmm... I thought the package had a trivial example where the standard
\code{ks.test()} was a problem?
\begin{Schunk}
\begin{Sinput}
> library(ks.test)
> x <- sample(1:10, 25, replace = TRUE)
> x
\end{Sinput}
\begin{Soutput}
 [1]  3  4  6 10  3  9 10  7  7  1  3  2  7  4  8  5  8 10
[19]  4  8 10  3  7  2  3
\end{Soutput}
\begin{Sinput}
> ks.test(x, ecdf(1:10))
\end{Sinput}
\begin{Soutput}
	One-sample Kolmogorov-Smirnov test

data:  x 
D = 0.08, p-value = 1
alternative hypothesis: two-sided 
\end{Soutput}
\begin{Sinput}
> ks.test(x, ecdf(1:10), alternative = "g")
\end{Sinput}
\begin{Soutput}
	One-sample Kolmogorov-Smirnov test

data:  x 
D^+ = 0.04, p-value = 0.7731
alternative hypothesis: the CDF of x lies above the null hypothesis 
\end{Soutput}
\begin{Sinput}
> ks.test(x, ecdf(1:10), alternative = "g", exact = FALSE)
\end{Sinput}
\begin{Soutput}
	One-sample Kolmogorov-Smirnov test

data:  x 
D^+ = 0.04, p-value = 0.9231
alternative hypothesis: the CDF of x lies above the null hypothesis 
\end{Soutput}
\end{Schunk}

Discuss this, nice example.  Relate to new material that is about to go into
the earlier section.
\begin{Schunk}
\begin{Sinput}
> stats::ks.test(c(0, 1), ecdf(c(0, 1)))
\end{Sinput}
\begin{Soutput}
	One-sample Kolmogorov-Smirnov test

data:  c(0, 1) 
D = 0.5, p-value = 0.5
alternative hypothesis: two-sided 
\end{Soutput}
\begin{Sinput}
> ks.test(c(0, 1), ecdf(c(0, 1)))
\end{Sinput}
\begin{Soutput}
	One-sample Kolmogorov-Smirnov test

data:  c(0, 1) 
D = 0, p-value = 1
alternative hypothesis: two-sided 
\end{Soutput}
\end{Schunk}

JAY: simulation showing original ks.test() was conservative, Conover exact.

\begin{Schunk}
\begin{Sinput}
> library(cvm.test)
> cvm.test(x, ecdf(1:10))
\end{Sinput}
\begin{Soutput}
	Cramer-von Mises - W2

data:  x 
W2 = 0.057, p-value = 0.8114
alternative hypothesis: Two.sided 
\end{Soutput}
\begin{Sinput}
> cvm.test(x, ecdf(1:10), type = "A2")
\end{Sinput}
\begin{Soutput}
	Cramer-von Mises - A2

data:  x 
A2 = 0.3969, p-value = 0.75
alternative hypothesis: Two.sided 
\end{Soutput}
\end{Schunk}

TAYLOR: good cyclical example, relating to new material about to go into
earlier section.

@@

Go into more detail about reasons for using these versus chi-squared type tests. (*)

Issues revolving around incorrect usage of default ks.test() in R because of continuity error (i.e. ks.test(1:2, ecdf(1:2)).  Will be earlier, needed here?

Further discuss numerical issues; ks -> n over 30 has problems, CvM -> integrate function can sometimes refuse to run because of failure to calculate integral.
Earlier, needed here?

Benefits and drawbacks of using simulation to gather p-values. (*)

Possibly go into a brief discussion of classes and class inheritance in R. The use of `htest' in our functions; benefits and issues with using it. Particularly troublesome in ks.test where we have a range of p-values.  More precisely, explain that we
could provide upper and lower bounds for the p-value but are limited by class
\code{htest}.

Two-sample cases; no theory for discrete distributions. Why? Distribution dependence messes things up. Discuss this for sure. (*)
